Proving Atomic Specifications for Data Structures with Fine-Grained Locking\thanks{Supported by the NSF (details).}

### 2022-04-08 concurrency/ASL 2022/intro.tex
restructured paper, updated todo

In this paper, we adapt the latter approach to the Verified Software Toolchain~\cite{plfcc}, a system for proving correctness of C programs that relies on style 1 lock specs for its soundness proof w.r.t. the CompCert compiler~\cite{compcert}, and has recently been extended with support for general ghost state and atomicity proofs. We demonstrate that 1) this approach can be used to prove atomic specs for interesting data structures with fine-grained locking, and 2) these proofs are much more complicated than those built with style 2 lock specs. The key difference is that in style 1, we must associate each locked component with some resources and then relate those resources to the state of the data structure as a whole, while in style 2 each lock access can interact directly with the full state of the data structure. However, the top-level specifications proved for the data structure are not affected by the complexity of the lock specs, so proofs of clients of the data structure are no more complex than they would be otherwise.

>>>>

In this paper, we adapt the latter approach to the Verified Software Toolchain~\cite{plfcc}, a system for proving correctness of C programs that relies on style 1 lock specs for its soundness proof w.r.t. the CompCert compiler~\cite{compcert}, and has recently been extended with support for general ghost state and atomicity proofs. We demonstrate that 1) \textbf{this approach can be used to prove atomic specs for interesting data structures with fine-grained locking}, and 2) \textbf{these proofs are much more complicated than those built with style 2 lock specs}. The key difference is that in style 1 we must associate each locked component with some resources and then relate those resources to the state of the data structure as a whole, while in style 2 each lock access can interact directly with the full state of the data structure. However, the top-level specifications proved for the data structure are not affected by the complexity of the lock specs, so proofs of clients of the data structure are no more complex than they would be otherwise. To our knowledge, this is the first formal comparison of the two styles of lock specifications, and our technique may be useful in tools like VST where the older style is deeply integrated into the tool.
